### 2019-10-31 paper/paper.tex
moar stuff for Lily :

First, some definitions.
The way I (Hogg) think about this is the following:
Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.
This is not precisely how the \expres\ team thinks about the problem,
I (Hogg) think.
For now I am going to ignore that fact, but I think what I say here
could be adapted to any \expres\ conventions.
Right now in the \expres\ pipeline the $\theta_{n}$ comprises the
amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
ignoring $y$?
If so, they have
\begin{equation}
\lambda = \sum_{i,j} c_{ij}\, x^i\,m^j
\quad ,
\end{equation}
where the $c_{ij}$ are coefficients.

If the space of all calibration possibilities is in fact
$K$-dimensional (where I am thinking of this as being a small integer,
like $K=3$ or $K=8$ or thereabouts), and if the calibration variations are so
small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
be replaced with a tiny model
\begin{equation}
\lambda = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions''
in calibration space that the spectrograph can depart from the
fiducial calibration.
The challenge is to learn these basis functions from the data, and get
the $K$ amplitudes $a_{nk}$ for every exposure $n$.

>>>>

First, some definitions.
The way I (Hogg) think about this is the following:
Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.
This is not precisely how the \expres\ team thinks about the problem,
I (Hogg) think.
For now I am going to ignore that fact, but I think what I say here
could be adapted to any \expres\ conventions.
Right now in the \expres\ pipeline the $\theta_{n}$ comprises the
amplitudes of some 9th-ish-order polynomial in $x$ and $n$, possibly
ignoring $y$?
If so, they (may?) have
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where the $c_{nij}$ are coefficients unique to exposure $n$, and
found for exposure $n$ without regard to any other exposure $n'$.
Further, I think they may be finding the coefficients $c_{nij}$ that
minimize an objective $Q$ that is something like the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is
treated as its own unique flower, I think there are improvements to be
made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.
If I am right about this, the fit could be taken to 13th order to obtain
much more expressiveness with (nearly) the same number of fit coefficients.
For another, the objective function could be made soft to permit
catastrophic outliers without destroying the fit.
I would recommend the iteratively reweighted least squares (IRLS).
This would make the fitting more robust.
For yet another, I have ideas about rescaling the products $x^i\,m^j$ to
protect the fitting from near-singularities or bad conditioning of the
linear-algebra operators.

Of course the main thing I want to do is go hierarchical!
In this case, the calibration of every image is informed by the
calibration of every other image.
The simplest form of hierarchical model is to use the calibration data
themselves to develop a low-dimensional basis for expressing the
calibration data.

If the space of all calibration possibilities is in fact
$K$-dimensional (where I am thinking of this as being a small integer,
like $K=3$ or $K=8$ or thereabouts), and if the calibration variations are so
small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions''
in calibration space that the spectrograph can depart from the
fiducial calibration.
The challenge is to learn these basis functions from the data, and get
the $K$ amplitudes $a_{nk}$ for every exposure $n$.

I know how to do this, but the margins here are too small to contain my
beautiful ideas...
